Author Guidelines for DAGM-OAGM Submission

### 2013-08-27 itm/paper.tex
describe use of dimension estimate in itm

A particularly drastic example is the case of the dataset having less samples
than features. In this case, the data clearly lies even on a linear subspace
of the input space, and the dimensionality of the input space does not
accurately reflect the intrinsic dimensionality of the data.
We use the estimate analyzed in \citet{massoud2007manifold}.
In their method, for each data point $x$, a local estimate $\hat{d}(x)$ of the
dimensionality at $x$ is computed as
\begin{equation}
    \hat{d}(x) = \frac{\ln 2}{\ln\left(r_k(x) / r_{\lfloor k /2 \rfloor}(x)\right)}.
\end{equation}
Here $k$ is a fixed integer and $r_k(x)$ is the distance of $x$ from its $k$th neighbor.
We follow \citet{massoud2007manifold} and set $k =\lceil 2 \ln n\rceil$.
The final estimate $\hat{d}$ is then computed by averaging the estimates over all $x$
\begin{equation}
    \hat{d} = \frac{1}{n}\sum_{x \in X} \min\left(\hat{d}(x),\ d\right).
\end{equation}
We found this estimate to work robustly and give sensible results for all
datasets we investigated.  As we already used a ball-tree data structure to
build the minimum spanning trees, estimating the dimensionality resulted only in
little computational overhead.

@@

\begin{table}[t]
\centering
\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lrrrccccc}
\toprule
\multicolumn{4}{c}{Dataset} &\multicolumn{4}{c}{Results}\\
\cmidrule{1-4}
\cmidrule{5-9}
Description &     n &    d &   k &  $k$-Means      & MeanNN        & SL            &   ITM        & ITM ID \\
\cmidrule{1-4}                   \cmidrule{5-5}    \cmidrule{6-6}  \cmidrule{7-7} \cmidrule{8-8} \cmidrule{9-9}
digits      &  1797 &   64 &  10 &     0.62 &      0.67            &0.10          &\B{0.85}       &0.73\\ 
faces       &   400 & 4096 &  40 &     0.41 &      0.49            &0.08          &0.02           &\B{0.54}\\ 
iris        &   150 &    4 &   3 &     0.72 &      0.75            &0.55          &\B{0.88}       &0.88\\
usps        &  9298 &  256 &  10 &     0.52 &      0.54            &0.00          &   0.44        &\B{0.64}\\ 
vehicle     &   846 &   18 &   4 &     0.10 &      0.09            &0.00          &\B{0.10}       &0.10\\
vowel       &   990 &   10 &  11 &     0.17 &      0.19            &0.00          &\B{0.20}       &0.19\\
waveform    &  5000 &   21 &   2 & \B{0.37} &      0.30            &0.00          &   0.23        &0.23\\
mnist       & 70000 &  784 &  10 &     0.37 &      N/A$^\dagger$         &0.00          &   0.50        &0.77\\
\bottomrule
\end{tabularx}
\caption{Adjusted Rand Index of $k$-Means, MeanNN, single link agglomerative clustering and
ITM on several datasets (higher is better). The best score for
each dataset is printed in bold.\\$^\dagger$\footnotesize We were unable to make MeanNN scale to 70000 data points, as storing the
whole pairwise distance matrix seems critical.\label{results}}
\end{table}

>>>>

A particularly drastic example is the case of the dataset having less samples
than features. In this case, the data clearly lies even on a linear subspace
of the input space, and the dimensionality of the input space does not
accurately reflect the intrinsic dimensionality of the data.
We use the estimate analyzed in \citet{massoud2007manifold}.
In their method, for each data point $x$, a local estimate $\hat{d}(x)$ of the
dimensionality at $x$ is computed as
\begin{equation}
    \hat{d}(x) = \frac{\ln 2}{\ln\left(r_k(x) / r_{\lfloor k /2 \rfloor}(x)\right)}.
\end{equation}
Here $k$ is a fixed integer and $r_k(x)$ is the distance of $x$ from its $k$th neighbor.
We follow \citet{massoud2007manifold} and set $k =\lceil 2 \ln n\rceil$.
The final estimate $\hat{d}$ is then computed by averaging the estimates over all $x$
\begin{equation}
    \hat{d} = \frac{1}{n}\sum_{x \in X} \min\left(\hat{d}(x),\ d\right).
\end{equation}
We compute the density estimate once, prior to clustering, and then plug the
estimate $\hat{d}$ into \eqref{graphobjective} in place of $d$.
We found this estimate to work robustly and give sensible results for all
datasets we investigated. As we already used a ball-tree data structure to
build the minimum spanning trees, we can reuse this structure to compute $r_k$.
Consequently, estimating the dimensionality resulted only in little
computational overhead.

@@

\begin{table}[t]
\centering
\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lrrrccccc}
\toprule
\multicolumn{4}{c}{Dataset} &\multicolumn{4}{c}{Results}\\
\cmidrule{1-4}
\cmidrule{5-9}
Description &     n &    d &   k &  $k$-Means      & MeanNN        & SL            &   ITM        & ITM ID \\
\cmidrule{1-4}                   \cmidrule{5-5}    \cmidrule{6-6}  \cmidrule{7-7} \cmidrule{8-8} \cmidrule{9-9}
digits      &  1797 &   64 &  10 &     0.62 &      0.67            &0.10          &\B{0.85}       &0.73\\ 
faces       &   400 & 4096 &  40 &     0.41 &      0.49            &0.08          &0.02           &\B{0.54}\\ 
iris        &   150 &    4 &   3 &     0.72 &      0.75            &0.55          &\B{0.88}       &0.88\\
usps        &  9298 &  256 &  10 &     0.52 &      0.54            &0.00          &   0.44        &\B{0.64}\\ 
vehicle     &   846 &   18 &   4 &     0.10 &      0.09            &0.00          &\B{0.10}       &0.10\\
vowel       &   990 &   10 &  11 &     0.17 &      0.19            &0.00          &\B{0.20}       &0.19\\
waveform    &  5000 &   21 &   2 & \B{0.37} &      0.30            &0.00          &   0.23        &0.23\\
mnist       & 70000 &  784 &  10 &     0.37 &      N/A$^\dagger$         &0.00          &   0.50        &0.77\\
\bottomrule
\end{tabularx}
\caption{Adjusted Rand Index of $k$-Means, MeanNN, single link agglomerative
    clustering and ITM on several datasets (higher is better). ITM ID refers to
    ITM using the estimated intrinsic dimensionality. The best score for each
    dataset is printed in bold.\\$^\dagger$\footnotesize We were unable to make
    MeanNN scale to 70000 data points, as storing the whole pairwise distance
    matrix seems critical.\label{results}}
\end{table}
